A Detection of the Environmental Dependence of the Stellar Profiles and Halos of Massive Central Galaxies

### 2017-07-10 draft/paper2.tex
Introduction edits I

    \jenny{Here is a suggested shorter introduction.}   
    One of the most intriguing observations of high-redshift galaxies from the last decade has 
    been the dramatic size growth in massive quiescent galaxies 
    \citep{Trujillo2006,vanDokkum2008,Damjanov2009,Cimatti2008} of factors of $2-4$ from $z \approx 2$ to the present. 
    These observations have led to a two-phase picture of galaxy growth,
    in which galaxies dissipatively form a central compact region and then their envelope grows around 
    them via dry merging \citep[e.g,][]{Naab2006, Khochfar2006,Oser2010, Oser2012}, which allows for 
    dramatic growth at late times. 
    There is likely also a contribution from progenitor bias, in which larger galaxies become quiescent 
    later \citep[e.g.,][]{Newman2012,XX}.

	There have been a number of papers testing the two-phase picture observationally at low redshift, 
    looking at the light profiles \citep{Huang2013a,Huang2013b,Oh2017}, color gradients (), 
    and stellar population gradients \citep[e.g.,][]{}. The observations are generically consistent with the 
    two-phase model. However, to really test this picture requires highly sensitive observations extending 
    beyond 10 kpc to truly probe the material accreted at late times \citep[Illustris paper here?][]{}. The surface 
    brightness falls rapidly in the outer parts of elliptical galaxies, making such observations very challenging. In 
    \citet[][; Paper I hereafter]{Huanetal2017}, we showed that the Hyper-Suprime Camera Survey {\bf CITE} allows us 
    to derive robust surface brightness profiles for {\it individual} galaxies out to 100 kpc.

    In Paper I, we exploited this sensitivity to verify that the fraction of accreted mass residing in an extended
    envelope is a function of the mass of the galaxy. Another unexplored issue is the extent to which, at fixed stellar 
    mass, the accreted fraction depends on the surrounding halo mass. According to this two-phase formation 
    scenario, the late time growth of massive ETGs is intrinsically tied to the growth of their host dark matter halos 
    (e.g. \citealt{Leauthaud2012, Behroozi2013, Shankar2013}). 
    Massive ETGs that live at the center of more massive halos should experience 
    a different mass assembly history and should be subject to a larger merger rate 
    compared to ETGs that live in smaller dark matter halos. 
    A natural prediction of the two-phase formation scenario is therefore that the 
    structural properties of massive galaxies should depend on their ``environment''. In this paper, we use our 
    large sample of early-type galaxies, along with halo masses from RedMapper \citep{}, to explore how the mass in
    extended stellar envelopes depends on halo mass. 
    Paper proceeds as follows: XXX.

    \song{This is the old version, needs work}
    Massive galaxies are unique laboratories for studying both cosmology and galaxy 
    formation. 
    Simulations of structure formation within the context of the $\Lambda$-CDM 
    cosmological model make predictions for the hierarchical growth of dark matter halos 
    and galaxies (e.g. \citealt{Baugh1996, DeLucia2006}), but there are many open 
    questions regarding the star-formation history, mass assembly process, and 
    structural evolution of massive galaxies. 
    Massive galaxies are thought to grow according to a ``two-phase'' formation scenario 
    (e.g. \citealt{Oser2010, Oser2012}). 
    In this scenario, the progenitors of $z{\sim} 0$ massive early-type galaxies (ETGs) 
    undergo a rapid growth phase at $z{\sim} 2$ triggered either by disk instabilities 
    or gas rich mergers (e.g. \citealt{Hopkins2008, Dekel2009}). 
    Observationally, the progenitors of ETGs are thought to correspond to the so-called 
    ``red nuggets'' (e.g. \citealt{Damjanov2009}) which have smaller effective radii 
    ($R_{\mathrm{e}}$; e.g. \citealt{Trujillo2006, vanDokkum2008, Cimatti2008}), 
    slightly higher central velocity dispersion and stellar mass density 
    (\mden{}; e.g. \citealt{vandeSande2011, Belli2014}), more disk-like morphologies 
    (e.g. \citealt{vanderWel2011}) than their local descendants 
    (e.g. \citealt{Bezanson2009, vanDokkum2010}).

    Following an initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quench star formation in 
    massive galaxies and a large fraction of these massive progenitors become quiescent 
    at $z>1$ (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase in the ``two-phase'' formation scenario is driven by 
    non-dissipative processes such as dry mergers with other galaxies 
    (e.g. \citealt{Naab2006, Khochfar2006}). 
    According to the ``two-phase'' scenario, minor mergers drive the growth of massive 
    galaxies at late times and can help explain the significant increase
    of $R_{\mathrm{e}}$ (e.g. \citealt{Newman2012, vdWel2014}) as well as the build-up 
    of the outer stellar halos of ETGs (e.g.  \citealt{Szomoru2012, Patel2013}).

    {\bf PP here about mass dependence: in the simulations, a larger fraction of the mass 
    is ex-situ at increasingly higher mass. Can we see evidence for that observationally by 
    looking at light profiles?}

    In addition, according to this two-phase formation scenario, the late time growth of massive ETGs 
    is intrinsically tied to the growth of their host dark matter halos 
    (e.g. \citealt{Leauthaud2012, Behroozi2013, Shankar2013}). 
    Massive ETGs that live at the center of more massive halos should experience 
    a different mass assembly history and should be subject to a larger merger rate 
    compared to ETGs that live in smaller dark matter halos. 
    A natural prediction of the two-phase formation scenario is therefore that the 
    structural properties of massive galaxies should depend on their ``environment''

    In order to truly understand the build-up of the outer parts of massive galaxies, very 
    deep data are needed. Massive ETGs have very steep inner stellar density profiles, 
    extremely extended stellar envelopes, and radial-dependent isophotal shapes which 
    make accurate measurements of \mstar{} and $R_{\mathrm{e}}$ very challenging
    (e.g. \citealt{Bernardi2013, DSouza2014}), 
    especially in the presence of shallow imaging data such as the Sloan Digital Sky 
    Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}).

    In this paper, we take advantage of new high-resolution (median 
    FWHM${\sim} 0.6^{\arcsec}$ in $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$\sb) imaging from the Hyper-Suprime Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to characterize the outer parts of massive galaxies 
    as a function of their properties. Here we will focus on (a) showing that we can 
    measure profiles for galaxies out to 100 kpc reliably with the HSC data and (b) looking at 
    trends in stellar mass and outer envelope properties. In a second paper in this series, we 
    will directly show the environmental dependence of the outer envelope mass.
    %The goal of this paper is to compare the \mden{} profiles and structural scaling 
    %relations of massive central galaxies in both low and high mass dark matter halos. 
    %Fig~\ref{fig:sdss_compare} displays a comparison of the imaging quality of massive 
    %ETGs using SDSS versus using the new HSC data. 
    The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. 
    With the help of the \redm{} cluster catalog (e.g.\ \citealt{Rykoff2014}; 
    \citealt{Rozo2015b}; please see Section~\ref{ssec:redmapper} for details), we 
    select a large sample (${\sim} 15000$) of massive central 
    galaxies with $0.3 < z < 0.5$ using ${\sim} 100$ square degree of data from the 
    HSC wide layer.

@@

    Magnitudes use the AB system (\citealt{Oke1983}), and are corrected for Galactic 
    extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_{\rm m}=0.3$, 
    and ${\Omega}_{\rm \Lambda}=0.7$.
    Stellar mass is noted \mstar{} and has been derived using a Chabrier Initial Mass 
    Function (IMF; \citealt{Chabrier2003}).     
    $M_{\mathrm{halo}}$ denotes halo mass in general whereas $M_{\rm 200b}$ is 
    explicitly defined as $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$).

@@

    %Previous work has shown that the half-light radii of massive ETGs is tightly 
    %correlated with \mstar{} (e.g. \citealt{Shen2003, Guo2009}). 
    %Models based on the ``two-phase'' formation scenario naturally predict that this 
    %mass--size relation should exhibit an environmental dependence: the effective radii 
    %of massive ETGs in more massive halos should be larger than those in less massive
    %halos at fixed \mstar{} (e.g. \citealt{Shankar2013, Shankar2014}).

    %However, while hints of an environmental dependence in the mass-size relation have 
    %been spotted at high-$z$ (e.g. \citealt{Papovich2012, Lani2013, Delaye2014}; but 
    %also see \citealt{Rettura2010}), there is no solid evidence for this effect at 
    %low-$z$ (e.g. \citealt{Nair2010, HCompany13}; but also see \citealt{Yoon2017}). 
    %The absence of an environmental dependence in the mass-size relation is an 
    %outstanding challenge to the two-phase formation scenario of ETGs.

>>>>

    One of the most intriguing observations of high-redshift galaxies from the last 
    decade has been the dramatic structural transformation in massive quiescent 
    galaxies \citep[e.g.][]{Trujillo2006, vanDokkum2008, Cimatti2008, Damjanov2009, 
    vanderWel2011, Szomoru2012, Patel2013} from $z \approx 2$ to the present. 
    These observations suggest that the progenitors of $z{\sim} 0$ massive early-type 
    galaxies (ETGs) need to increase their effective radius ($R_{\rm e}$) by factor 
    of 2--4 in the last 10 Gyrs (e.g. \citealt{Newman2012, vdWel2014}).
    They have led to a ``two-phase'' scenario for the formation of massive ETGs 
    (e.g. \citealt{Oser2010, Oser2012}),
    in which galaxies form a compact central region through highly dissipative 
    processes (e.g. gas-rich merger or cold gas-accretion; 
    \citealt{Hopkins2008, Dekel2009}) and then assemble their extended stellar 
    haloes via dry mergers 
    \citep[e.g,][]{Naab2006, Khochfar2006, Oser2010, Oser2012}, which allows for 
    significant growth at late times. 
    Meanwhile, progenitor bias, which proposes that larger ETGs were quenched more 
    recently, could also contribute to the apparent size growth, although its 
    important is under argument (e.g. \citealt{Newman2012, Carollo2013, 
    Poggianti2013, Belli2015, Keating2015, Fagioli2016}).

    Several previous works have tested the two-phase formation scenario 
    observationally using galaxies at low redshift, looking at their surface 
    brightness of mass density profiles (e.g. \citealt{Huang2013a, Huang2013b, 
    Oh2017}, optical color gradients (e.g. \citealt{LaBarbera2010, LaBarbera2012}), 
    and stellar population gradients \citep[e.g.,][]{Coccato2010, Coccato2011, 
    Greene2015, Barbosa2016}. 
    These observations are generally consistent with the two-phase scenario, 
    but this picture still face many challenges.
    One of the most outstanding issues is the extent to which, at fixed stellar mass, 
    the accreted fraction depends on the surrounding dark matter halo mass.

    According to this two-phase formation scenario, the late time growth of massive 
    ETGs is intrinsically tied to the growth of their host dark matter halos 
    (e.g. \citealt{Leauthaud2012, Behroozi2013, Shankar2013}). 
    Massive ETGs that live at the center of more massive halos should experience 
    different assembly and should be subject to higher merger rate compared to 
    ETGs with similar \mstar{} that live in less massive dark matter halos. 
    A natural prediction of this picture is therefore that the structures of 
    massive ETGs should depend on their ``environment''\footnote{There are 
    different definitions of ``environment'' in the literature.  
    In this work, we use ``environment'' and halo mass interchangeably.}.
    For instance, models based on two-phase scenario predict that the well-know 
    stellar mass--effective radius relation for ETGs (\mstar{}--$R_{\rm e}$ relation; 
    e.g., \citealt{Shen2003, Guo2009}) should show environmental dependence: the 
    $R_{\rm e}$ of massive ETGs in more massive halos should be larger than those 
    in less massive halos at fixed \mstar{} 
    (e.g. \citealt{Shankar2013, Shankar2014}).
    Yet, solid evidence of this effect at low redshift is still missing 
    (e.g. \citealt{Nair2010, HCompany13}; but also see \citealt{Yoon2017}).

    %However, while hints of an environmental dependence in the mass-size relation have 
    %been spotted at high-$z$ (e.g. \citealt{Papovich2012, Lani2013, Delaye2014}; but 
    %also see \citealt{Rettura2010}).

	% Song: Actually the Illustris paper suggests that, for massive galaxies, 
	%       even the inner region is dominated by accreted stars.

    To seriously test the environmental dependence of structure in massive ETGs 
    requires highly sensitive observations extending beyond the inner region of these 
    galaxies to truly probe the material accreted at late times. 
    This is very challenging observation as the surface brightness drops rapidly 
    with radius in massive ETGs.

    In \citet[][; Paper I hereafter]{hscMassiveI}, we showed that the deep, 
    multi-band images from the Subaru Strategic Program (SSP; \citealt{HSC_SSP,
    HSC_DR1}) using Hyper Suprime-Cam (HSC; \citealt{Miyazaki2012}, 
    Miyazaki in~prep.) allows us to extract robust surface stellar
    mass density (\mden{}) profiles for {\it individual} massive galaxies 
    at $0.3 < z < 0.5$ out to 100 kpc. 
    In Paper I, we exploited this sensitivity to show that the \mden{} profiles 
    and the shapes of the stellar halos of massive ETGs has a clear \mstar{} 
    dependence. 
    And the fraction of accreted stars residing in the extended envelope increases
    with the total \mstar{} of the galaxy.

    In this paper, we explore how the extended stellar halo in massive central galaxies
    depends on dark matter halo mass with the help of HSC images of a large sample 
    of massive ETGs and the \redm{} cluster catalog (e.g.\ \citealt{Rykoff2014}; 
    \citealt{Rozo2015b}; please see Section~\ref{ssec:redmapper} for details)

@@

    Magnitudes use the AB system (\citealt{Oke1983}), and are corrected for Galactic 
    extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_{\rm m}=0.3$, 
    and ${\Omega}_{\rm \Lambda}=0.7$.
    Stellar mass is denoted \mstar{} and has been derived using a Chabrier Initial Mass 
    Function (IMF; \citealt{Chabrier2003}).

    Halo mass is defined as $M_{\rm 200b}$ as 
    $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$).

@@
